---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

```{r}

# Your answer here.

# num 1 and 4

```

2E2.

```{r}

# Your answer here.
# num 3
```

2E3.

```{r}

# Your answer here.
# 1 and 4
```

2E4.

```{r}

# Discuss in pairs/groups. Ask me for a hint if you feel like it.

#we dont konw the excact mechanisms between the throw

```

### Medium

2M1.

```{r}

# Write your code here.
p_grid = seq(from = 0, to = 1, length.out=20)

prior = rep(1, 20)

# x is the number of successes and size is the number of times the globe is thrown
likelihood1 = dbinom(3, size = 3, prob = p_grid)
likelihood2 = dbinom(3, size = 4, prob = p_grid)
likelihood3 = dbinom(5, size = 7, prob = p_grid)
#so this is the probability of hitting this amount of water (W) given the different probabilities of hitting water (0-100%) which is the likelihood.
# but we want the reverse, eg. the probability of hitting water the the globe given the samples we got.


print(likelihood1)
#print(likelihood2)
#print(likelihood3)

# compute product of likelihood and prior (the model)
unstd.posterior1 = likelihood1 * prior
unstd.posterior2 = likelihood2 * prior
unstd.posterior3 = likelihood3 * prior
print(unstd.posterior1) #should be the same as likely hood since prior is uniform

# standardize the posterior so it sums to 1
posterior1 = unstd.posterior1 / sum(unstd.posterior1) # this is the likelihood out of all the likelyhoods.
posterior2 = unstd.posterior2 / sum(unstd.posterior2)
posterior3 = unstd.posterior3 / sum(unstd.posterior3)

# plot the distribution
plot(p_grid, posterior1, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
plot(p_grid, posterior2, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
plot(p_grid, posterior3, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

2M2.

```{r}

# In the book, McElreath says that we can do much better than choosing a uniform prior for estimating the coverage of water. Is this what we are doing here? How does that impact our estimation?

# Write your code here.
p_grid = seq(from = 0, to = 1, length.out=20)

# over the 20 sequences between 0 and 100% prob we set prior to 100% after the first ten in order to show that there was at least 50% water in
# the other planets where this was tested.
prior = c(0:9*0, rep(1, 10))

# x is the number of "successes" and size is the number of times the globe is thrown
likelihood1 = dbinom(3, size = 3, prob = p_grid)
likelihood2 = dbinom(3, size = 4, prob = p_grid)
likelihood3 = dbinom(5, size = 7, prob = p_grid)
#so this is the probability of hitting this amount of water (W) given the different probabilities of hitting water (0-100%) which is the likelihood.
# but we want the reverse, eg. the probability of hitting water the the globe given the samples we got.


#print(likelihood1)
#print(likelihood2)
#print(likelihood3)

# compute product of likelihood and prior
unstd.posterior1 = likelihood1 * prior
unstd.posterior2 = likelihood2 * prior
unstd.posterior3 = likelihood3 * prior

# standardize the posterior so it sums to 1
posterior1 = unstd.posterior1 / sum(unstd.posterior1)
posterior2 = unstd.posterior2 / sum(unstd.posterior2)
posterior3 = unstd.posterior3 / sum(unstd.posterior3)

# plot the distribution
plot(p_grid, posterior1, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
plot(p_grid, posterior2, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")
plot(p_grid, posterior3, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

2M3.

```{r}
# Use the Bayes formula!

mars = dbinom(1, size = 1, prob = 1)
earth = dbinom(1, size = 1, prob = 0.7)

#likelyhood * prior / the probability of hitting water in all scenarios.

prob_mars = (1*0.5)/0.85
prob_earth = (0.7*0.5)/0.85

cat("mars", prob_mars)
print("")
cat("earth", prob_earth)

```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4.

```{r}

# Write your probability here.

```

2M5.

```{r}

# Write your probability here.

```

2M6.

```{r}

# Write your probability here.

```

2M7
```{r}
#there is 2/3 for the card being black on the other side: if it was black the first card could give that outcome 0% then 50% and 100% of the time. then we need to normalize it. there is 50/150 it could be the black and white and 100/150% it could be the black/black.

#when the white card is then drawn there is 3 possibilities, 0% the black/black could yield this, 100% the white white could yield this, and 50% the black/white could yield this. Now there is a 33% chance that the black/white card has already been picked from before and therefore 33% less chance of the 1/3 chance there was before. therefore we take 1/3 of the 50% probability that the black/white card could yield white = 33%. Therefore now the accumulated probability is sjull100+33 and the probability of picking the white/black card is 33/133= 25%. therefore the 33% chance that the first card was black before decreases by 25% so the probability that it was the black/white card is now 1/3-1/3*25% = 1/4. Therefore the rest that is left is 75% for the black/black card.

```

