---
title: "Chapter 3"
output: html_document
date: "2024-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(100)
```

# Exercises for Class 2

What's good my young Bayesians? Welcome to the exercises notebook for your 2nd class of Methods 4. As you might have noticed, a lot of Chapter 3 deals with summarizing the posterior. Summarizing the posterior is helpful, but as both Chris and Richard have re-iterated multiple times, you should **always report the full posterior distribution**.

Therefore, I am providing the solutions to the *Easy* exercises aimed at summarizing the posterior, so you don't waste time scanning the book for the right lines of code. Play around with them, but don't wait too long to get to the **juice** of this notebook (and Chapter 3) - **simulation of the model's implied observations**.

## Exercises

### Easy.

These problems use the samples form the posterior distribution for the globe tossing example. This code will give you a specific set of samples, so that you can check your answers correctly.

```{r}

p_grid = seq(from = 0, to = 1, length.out=1000)
prior = rep(1, 1000)
likelihood = dbinom(6, size = 9, prob = p_grid)
posterior = likelihood * prior
posterior = posterior / sum(posterior)
samples = sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)

# Let's also visualize our samples, just to know what we are working with
dens(samples)

```

#### 3E1. 

How much posterior probability lies below p = 0.2?

```{r}

sum(samples < 0.2) / 10000

```

#### 3E2. 

How much posterior probability lies above p = 0.8?

```{r}

sum(samples > 0.8) / 10000


```

#### 3E3. 

How much posterior probability lies above p = 0.2 and p = 0.8?

```{r}

sum(samples > 0.2 & samples < 0.8) / 10000

```

#### 3E4. 

20% of the posterior probability lies below which value of p?

```{r}

quantile(samples, 0.2)

```

#### 3E5. 

20% of the posterior probability lies above which value of p?

```{r}

quantile(samples, 0.8)

```

#### 3E6. 

Which values of p contain the narrowest interval equal to 66% of the posterior probability?

```{r}

HPDI(samples, prob = 0.66)

```

#### 3E7. 

Which values of p contain the 66% of the posterior probability, assuming equal posterior probability both below and above the interval?

```{r}

PI(samples, prob = 0.66)

```

### Sampling to simulate prediction

Section **3.3** of this Chapter is, imo, the most important one. Simulating predictions is something you'll be frequently doing going forward, so it's worth taking time to unpack the mechanics of it.

```{r}
# First, play around with this function. What does it do? What does the output mean?
rbinom(1, size = 9, prob = 0.7)

# Now let's scale it up.
dummy_w = rbinom(1e4, size = 9, prob = 0.1)
simplehist(dummy_w)

# What happens if we increase the size?
# What about probability?
```

Now let's do a posterior predictive check. Take a minute to contemplate what are we doing here. How does the resulting distribution compare to our sampled posterior?

```{r}
w = rbinom(1e4, size = 9, prob = samples)
simplehist(w)

# Compare to sampled posterior here - choose one of the summarization methods above! Is it more or less conservative?
```

The resulting distribution is for predictions, but it incorporates all of the uncertainty embodied in the posterior distribution for the parameter p. As a result, it is honest. While the model does a good job of predicting the data - the most likely observation is indeed the observed data - predictions are still quite spread out.

### Medium 

### 3M1 & 3M2. 

Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as before.

Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HDPI for p.

```{r}

p_grid = seq(from = 0, to = 1, length.out=1000)
prior = rep(1, 1000)
likelihood = dbinom(8, size = 15, prob = p_grid)
posterior = likelihood * prior
posterior = posterior / sum(posterior)
samples = sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)

dens(samples)

```

```{r}

# Your code here.

```

#### 3M3.

Construct a posterior predictive check for this model and data. This means simulate the distribution of samples, averaging over the posterior uncertainty in p. What is the probability of observing 8 water in 15 tosses?

```{r}

# Your code here.

```

#### 3M4. 

Using the posterior distribution constructed from the new (8/15) data, now calculate the probability of observing 6 water in 9 tosses.

```{r}

# Your code here.

```

#### 3M5.

Start over at 3M1, but now use a prior that is zero below p=0.5 and a constant above p=0.5. This corresponds to prior information that a majority of the Earth’s surface is water. Repeat each problem above and compare the inferences. What difference does the better prior make? If it helps, compare inferences (using both priors) to the true value p = 0.7.

```{r}

# Your code here.

```


### Hard

**Introduction**. The practice problems here use the data below. These data indicate the gender (male = 1, female = 0) of officially reported first and second born children in 100 two-children families.

```{r}
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)
```

So, for example, the first family in the data reported a boy (1) and then a girl (0). The second family reported a girl (0) and then a boy (1). The third family reported two girls. You can load these wo vectors into R's memory by typing:

```{r}

#library(rethinking)
data(homeworkch3)

```

#### 3H1. 

Using grid approximation, compute the posterior distribution for the probability of a birth being a boy. Assume a uniform prior probability. Which parameter value maximizes the posterior probability?

```{r}

# Your code here.

```

#### 3H2. 

Using the sample function, draw 10,000 random parameter values from the posterior distribution you calculated above. Use these samples to estimate the 50%, 89% and 97% highest posterior density intervals.

```{r}

# Your code here. 

```

#### 3H3. 

Use rbinom to simulate 10,000 replicates of 200 births. You should end up with 10,000 numbers, each one a count of boys out of 200 births. Compare the distribution of predicted numbers of boys to the actual count in the data (111 boys out of 200 births). There are many good ways to visualize the simulations, but the dens command (part of rethinking package) is probably the easiest way in this case. Does it look like the model fits the data well? That is, does the distribution of the predictions include the actual observation as a central, likely outcome?

```{r}

# Your code here.

```
